Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition

### 2011-02-17 paper/main.tex
Moved Allen's additional defs to mydefs.sty
Reflowed long lines

\def\arr{{\mathcal{A}}}
\newcommand{\Null}{{\mbox{Null}}}
\renewcommand{\mathbf}{\boldsymbol}
\renewcommand{\Re}{{\mathbb{R}}}
\def\tt{{\boldsymbol{t}}}
\def\ee{{\boldsymbol{e}}}
\def\aa{{\boldsymbol{a}}}
\def\bb{{\boldsymbol{b}}}
\def\dd{{\boldsymbol{d}}}
\def\ss{{\boldsymbol{s}}}
\def\cc{{\boldsymbol{c}}}
\def\pp{{\boldsymbol{p}}}
\def\chat{{\boldsymbol{\hat{\c}}}}
\def\nn{{\boldsymbol{n}}}
\def\rr{{\boldsymbol{n}}}
\def\m{{\boldsymbol{m}}}
\def\X{{\boldsymbol{X}}}
\def\L{{\boldsymbol{L}}}
\def\l{{\boldsymbol{l}}}
\def\Y{{\boldsymbol{Y}}}
\def\V{{\boldsymbol{V}}}
\def\uu{{\boldsymbol{u}}}
\def\vv{{\boldsymbol{v}}}
\def\yy{{\boldsymbol{y}}}
\def\xx{{\boldsymbol{x}}}
\def\ww{{\boldsymbol{w}}}
\def\zz{{\boldsymbol{z}}}
\def\uu{{\boldsymbol{u}}}
\def\F{{\boldsymbol{F}}}
\def\veronese{{\nu}}
\def\Qveronese{{\mu}}
\def\H{{\mathcal{H}}}
\def\Q{{\mathcal{Q}}}
\def\E{{\mathcal{E}}}
\def\L{{\mathcal{L}}}
\def\M{{\mathcal{M}}}
\def\A{{\mathcal{A}}}
\def\B{{\mathcal{B}}}
\def\F{{\mathcal{F}}}
\def\field{\mathbb{F}}
\def\QED{~\rule[-1pt]{5pt}{5pt}\par\medskip}

@@

\section{Introduction}

$\ell_1$-minimization ($\ell_1$-min) has received much attention in recent years due to important applications
in compressive sensing \cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.
The problem refers to finding the minimum $\ell$-norm solution to an underdetermined linear system $\bb=A\xx$:
\begin{equation}
\min \|\xx\|_1\quad \mbox{ subj. to }\quad \bb = A\xx.
\label{eq:l1min}
\end{equation}
It has now been well understood that, under certain conditions \cite{CandesE2005-IT_1,DonohoD2004}, the minimum $\ell_1$-norm solution is
also the \emph{sparsest} solution to the system \eqref{eq:l1min}.

@@

%\subsection{Contributions of the Paper}

@@

Note that in the iterative procedure outlined in \eqref{eq:alm}, the second step only involves vector algebra and matrix-vector multiplication. Therefore, the procedure is computationally efficient only if it is easier to minimize the augmented Lagrange function compared to solving the original problem \eqref{eq:l1min} directly. In fact, the first step that minimizes $L_{\mu_k} (\xx,\yy_k)$ assuming $\yy_k$ and $\mu_k$ are fixed can be solved element-wise iteratively by a soft-thresholding algorithm \cite{WrightS2008,BeckA2009}, whose time complexity is bounded by $O(n^2)$ and can be easily parallelized. Algorithm \ref{alg:alm} summarizes the ALM $\ell_1$-min algorithm.
 \begin{algorithm}[h]
\caption{Augmented Lagrange Multiplier (ALM)}
{\bf INPUT:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$, $\tau\leftarrow$ largest eigenvalue of $A^TA$, and a constant $\rho>1$.
\begin{algorithmic}[1]
\WHILE{not converged ($k = 1,2,\ldots$)} \STATE $t_1 \leftarrow 1$, $\zz_1
\leftarrow \xx_k$, $\uu_1 \leftarrow \xx_k$ \WHILE{not
converged ($l = 1,2,\ldots$)} \STATE $\uu_{l+1}  \leftarrow
\mathrm{soft}\left(\zz_l - \frac{1}{\tau}A^T\left(A\zz_l - \bb
- \frac{1}{\mu_k}\yy_k\right), \frac{1}{\mu_k\tau}\right)$
\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
\sqrt{1+4t_l^2}\right)$ \STATE $\zz_{l+1} \leftarrow \uu_{l+1}
+ \frac{t_l - 1}{t_{l+1}}(\uu_{l+1} - \uu_l)$ \ENDWHILE \STATE$
\xx_{k+1} \leftarrow \uu_{l+1}$ \STATE $\yy_{k+1} \leftarrow
\yy_k + \mu_k (\bb - A\xx_{k+1})$ \STATE $\mu_{k+1} \leftarrow
\rho\cdot\mu_k$ \ENDWHILE 
\end{algorithmic}

@@

\subsection{GPU Hardware Parallelism}
%High-level description of GPU parallelism, i.e.

@@

% This section is Drew's
\section{Parallelization of the Face Alignment Stage}
\label{sec:alignment}
{\bf AY: Add an algorithm here describe the alignment block in the pipeline. Explain the algorithm and necessary modification from generic l-1}

>>>>

@@

\section{Introduction} 
$\ell_1$-minimization ($\ell_1$-min) has received much attention in recent
years due to important applications in compressive sensing
\cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.  The
problem refers to finding the minimum $\ell$-norm solution to an
underdetermined linear system $\bb=A\xx$:
\begin{equation}
\min \|\xx\|_1\quad \mbox{ subj. to }\quad \bb = A\xx.
\label{eq:l1min}
\end{equation}
It has now been well understood that, under certain conditions
\cite{CandesE2005-IT_1,DonohoD2004}, the minimum $\ell_1$-norm solution is also
the \emph{sparsest} solution to the system \eqref{eq:l1min}.

@@

@@

Note that in the iterative procedure outlined in \eqref{eq:alm}, the second
step only involves vector algebra and matrix-vector multiplication. Therefore,
the procedure is computationally efficient only if it is easier to minimize the
augmented Lagrange function compared to solving the original problem
\eqref{eq:l1min} directly. In fact, the first step that minimizes $L_{\mu_k}
(\xx,\yy_k)$ assuming $\yy_k$ and $\mu_k$ are fixed can be solved element-wise
iteratively by a soft-thresholding algorithm \cite{WrightS2008,BeckA2009},
whose time complexity is bounded by $O(n^2)$ and can be easily parallelized.
Algorithm \ref{alg:alm} summarizes the ALM $\ell_1$-min algorithm.

 \begin{algorithm}[h]
\caption{Augmented Lagrange Multiplier (ALM)}
{\bf INPUT:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$, $\tau\leftarrow$ largest eigenvalue of $A^TA$, and a constant $\rho>1$.
\begin{algorithmic}[1]
\WHILE{not converged ($k = 1,2,\ldots$)} \STATE $t_1 \leftarrow 1$, $\zz_1
\leftarrow \xx_k$, $\uu_1 \leftarrow \xx_k$ \WHILE{not
converged ($l = 1,2,\ldots$)} \STATE $\uu_{l+1}  \leftarrow
\mathrm{soft}\left(\zz_l - \frac{1}{\tau}A^T\left(A\zz_l - \bb
- \frac{1}{\mu_k}\yy_k\right), \frac{1}{\mu_k\tau}\right)$
\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
\sqrt{1+4t_l^2}\right)$ \STATE $\zz_{l+1} \leftarrow \uu_{l+1}
+ \frac{t_l - 1}{t_{l+1}}(\uu_{l+1} - \uu_l)$ \ENDWHILE \STATE$
\xx_{k+1} \leftarrow \uu_{l+1}$ \STATE $\yy_{k+1} \leftarrow
\yy_k + \mu_k (\bb - A\xx_{k+1})$ \STATE $\mu_{k+1} \leftarrow
\rho\cdot\mu_k$ \ENDWHILE 
\end{algorithmic}

@@

\subsection{GPU Hardware Parallelism}

@@

\section{Parallelization of the Face Alignment Stage}
\label{sec:alignment}
{\bf AY: Add an algorithm here describe the alignment block in the pipeline. Explain the algorithm and necessary modification from generic l-1}
